# pdfsearch

[![Build Status](https://travis-ci.org/lebebr01/pdfsearch.svg?branch=master)](https://travis-ci.org/lebebr01/pdfsearch)
[![codecov.io](https://codecov.io/github/lebebr01/pdfsearch/coverage.svg?branch=master)](https://codecov.io/github/lebebr01/pdfsearch?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/pdfsearch)](http://cran.r-project.org/package=pdfsearch)

This package defines a few useful functions for keyword searching using the [pdftools](https://github.com/ropensci/pdftools)  package developed by [rOpenSci](https://ropensci.org/).

The package can be installed from CRAN directly:
```{r install_cran, eval = FALSE}
install.packages("pdfsearch")
```

To install the development version you use devtools:
```{r install, eval = FALSE}
install.packages("devtools")
devtools::install_github('lebebr01/pdfsearch')
```

## Basic Usage
There are currently two functions in this package of use to users. The first `keyword_search` takes a single pdf and searches for keywords from the pdf. The second `keyword_directory` does the same search over a directory of pdfs.

## Example with `keyword_search`
The package comes with two pdf files from [arXiv](https://arxiv.org/) to use as test cases. Below is an example of using the `keyword_search` function.
```{r search1, message = FALSE}
library(pdfsearch)
file <- system.file('pdf', '1610.00147.pdf', package = 'pdfsearch')

result <- keyword_search(file, 
            keyword = c('measurement', 'error'),
            path = TRUE)
head(result)
head(result$line_text, n = 2)
```

The location of the keyword match, including page number and line number, the actual line of text, and a tokenized version of the text (raw text split by individual words) are returned by default.

In addition, by default the hyphenated words at the end of the text are combined with the continued word at the start of the next line. If this behavior is not of interest, set the `remove_hyphen` argument to `FALSE`.

### Surrounding lines of text 
It may be useful to extract not just the line of text that the keyword is found in, but also surrounding text to have additional context when looking at the keyword results. This can be added by using the argument `surround_lines` as follows:
```{r surround}
file <- system.file('pdf', '1610.00147.pdf', package = 'pdfsearch')

result <- keyword_search(file, 
            keyword = c('measurement', 'error'),
            path = TRUE, surround_lines = 1)
head(result)
head(result$line_text, n = 2)
```

## Example with `keyword_directory`
The `keyword_directory` function allows users to search for keywords in multiple PDF files in one function call. The same functionality from the `keyword_search` function can be invoked, specifically `remove_hyphen` and `surround_lines`. Below is an example of searching a single directory. 

```{r directory}
directory <- system.file('pdf', package = 'pdfsearch')

# do search over two files
directory_result <- keyword_directory(directory, 
       keyword = c('repeated measures', 'measurement error'),
       surround_lines = 1, full_names = TRUE)

head(directory_result, n = 2)
```

Note here the use of the `full_names` argument. This argument specifies whether the full path to the PDF file should be used (`TRUE`) or if only the file name should be used (`FALSE`; default). A few other useful arguments are possible when searching for keywords within multiple PDF files in a directory. One is the `recursive` (default is `FALSE`), where if set to `TRUE` will search within subdirectories as well, the default function behavior will not venture into subdirectories. Finally, if the directory has many PDF files, testing the function first on a handful of PDF files may be desired. The number of PDF files can be limited with the argument `max_search` where a positive integer can be specified indicating the number of PDF files to search. For example, is `max_search = 2`, only the first two PDF files will be searched within the directory.

### Shiny App
The package also has a simple Shiny app that can be called using the following command
```{r shiny, eval = FALSE}
run_shiny()
```


